---
title: 'Inter-clonal competition over queen succession imposes a cost of parthenogenesis on termite colonies'
output:
  html_document: default
  pdf_document: default
---

# packages
```{r}
install.packages("pacman")
pacman::p_load(
     readxl,
     ggplot2,
     tidyverse,
     dplyr,
     stringr,
     lme4,
     car,
     epitools,
     survival,
     survminer,
     scales)
```

# Distribution of the number of secondary queens per colony
## * data loading
```{r}
Fig2a <- read_excel("Dataset S1.xlsx", sheet="no. of royals") %>% 
  as.data.frame(.)
```

## * sample size
```{r}
summary_Fig2a <- Fig2a %>% 
  group_by(no_of_primary_queen) %>% 
  summarize(number = n())
summary_Fig2a
```

## * statistical analysis
```{r}
model1 <- glm(no_of_secondary_queen ~ as.factor(no_of_primary_queen), family = poisson, Fig2a)
model2 <- glm(no_of_secondary_queen ~ 1, family = poisson, Fig2a)
summary(model1)
Anova(model1)
anova(model1, model2, test="Chi")
```

## * figure
```{r}
xmin <- 0 
xmax <- 320 
xscale <- 50 
binwidth <- 10 
ymin <- 0 
ymax <- 33 
yscale <- 10 

par(ps=18)

Fig_2a <- 
  ggplot(Fig2a, aes(x = no_of_secondary_queen, fill = as.factor(no_of_primary_queen))) +
  geom_histogram(boundary = 0, 
                 binwidth = binwidth, 
                 closed ='left' , 
                 alpha = 1, 
                 color = "black", 
                 position = "identity" 
                 ) +
  scale_fill_manual(values = c("white", "black")) + 
  scale_x_continuous(breaks = seq(xmin, xmax, by = xscale), lim = c(xmin, xmax), expand = c(0,0)) + 
  scale_y_continuous(breaks = seq(ymin, ymax, by = yscale), lim = c(ymin, ymax), expand = c(0,0)) + 
  labs(y = "no. of colonies", 
       x = "no. of secondary queens / colony",
       title ="") +
  theme_classic()
Fig_2a

# save figure
# ggsave(file = "Figure 2a.pdf", plot = Fig_2a, dpi = 300, width = 140, height = 89, units = c("mm"))
```

# Relationship between total weight of secondary queens in a colony and the number of clone types
## * data loading
```{r}
Fig2bc <- read_excel("Dataset S1.xlsx", sheet="no. of clones in SQ population") %>% 
  as.data.frame(.)
```

## * sample size
```{r}
summary_Fig2bc <- Fig2bc %>% 
  group_by(no_of_primary_queen) %>% 
  summarize(number = n())
summary_Fig2bc
```

## * statistical analysis
### * comparison of number of clones between colonies with and without PQ
```{r}
model_cdq1 <- glmer(cbind(no_of_clone_types,no_of_secondary_queens_genotyped-no_of_clone_types) ~ as.factor(no_of_primary_queen) + (1|colony), family = binomial, Fig2bc)
model_cdq2 <- glmer(cbind(no_of_clone_types,no_of_secondary_queens_genotyped-no_of_clone_types) ~ 1 + (1|colony), family = binomial, Fig2bc)
summary(model_cdq1)
Anova(model_cdq1)
anova(model_cdq1, model_cdq2, test="Chi")
```

### * relationship between colony size and number of clones in SQ population
```{r}
model_cd1 <- glmer(cbind(no_of_clone_types,no_of_secondary_queens_genotyped-no_of_clone_types) ~ total_weight_of_queen + (1|colony), family = binomial, Fig2bc)
model_cd2 <- glmer(cbind(no_of_clone_types,no_of_secondary_queens_genotyped-no_of_clone_types) ~ 1 + (1|colony), family = binomial, Fig2bc)
summary(model_cd1)
Anova(model_cd1)
anova(model_cd1, model_cd2, test="Chi")
```

## * figure
```{r}
Fig_2b <- 
  ggplot(Fig2bc, aes(y = no_of_clone_types, 
                     x = total_weight_of_queen, 
                     fill = as.factor(no_of_primary_queen))) +
  stat_function(fun = function(x){(1/(1+exp(-(summary(model_cd1)[[10]][1] + summary(model_cd1)[[10]][2]*x))))*20}, color = "black") +
  geom_point(size = 2, stroke = 1, shape=21) +
  scale_fill_manual(values = c("white", "black")) +
  scale_y_continuous(breaks=seq(0,20,by=5),limits=c(0,20),expand=c(0,0)) +
  scale_x_continuous(breaks=seq(0,3000,by=200),limits=c(0,1100),expand=c(0,0)) +
  theme_classic() +
  theme(legend.justification = c(1, 1),　
        legend.text = element_text(size=8), 
        legend.title = element_text(size=8), 
        axis.title.x = element_text(size=8), 
        axis.title.y = element_text(vjust = 0.5,angle=90,size=8), 
        axis.text.x = element_text(size=8,color="black"), 
        axis.text.y = element_text(size=8,color="black"),
        axis.ticks.x = element_line(size = 0.5), 
        axis.line = element_line(size = 0.5, 
                                linetype = "solid", 
                                lineend = "round")
        ) +
  labs(y = "no. of clone groups", 
       x = "total weight of queens (mg)",
       title ="")
Fig_2b

# save data
# ggsave(file = "Figure 2b.pdf", plot = Fig_2b, dpi = 300, width = 140, height = 89, units = c("mm"))
```

# Relationship between colony size and the proportion of the most dominant clone
## * statistical analysis
### * comparison of the proportion of the most dominant clone between colonies with and without PQ
```{r}
model_cdq1 <- glmer(cbind(no_of_individuals_of_the_most_dominate_clone_type,no_of_secondary_queens_genotyped-no_of_individuals_of_the_most_dominate_clone_type) ~ total_weight_of_queen + (1|colony), family = binomial, Fig2bc)
model_cdq2 <- glmer(cbind(no_of_individuals_of_the_most_dominate_clone_type,no_of_secondary_queens_genotyped-no_of_individuals_of_the_most_dominate_clone_type) ~ 1 + (1|colony), family = binomial, Fig2bc)
summary(model_cdq1)
Anova(model_cdq1)
anova(model_cdq1, model_cdq2, test="Chi")
```

### * relationship between colony size and the proportion of the most dominant clone in SQ population
```{r}
model_cd1 <- glmer(cbind(no_of_individuals_of_the_most_dominate_clone_type,no_of_secondary_queens_genotyped-no_of_individuals_of_the_most_dominate_clone_type) ~ total_weight_of_queen + (1|colony), family = binomial, Fig2bc)
model_cd2 <- glmer(cbind(no_of_individuals_of_the_most_dominate_clone_type,no_of_secondary_queens_genotyped-no_of_individuals_of_the_most_dominate_clone_type) ~ 1 + (1|colony), family = binomial, Fig2bc)
summary(model_cd1)
Anova(model_cd1)
anova(model_cd1, model_cd2, test="Chi")
```

## * figure
```{r}
Fig_2c <- 
  ggplot(Fig2bc, aes(y = percent_of_the_most_domonate_clone_type/100, 
                     x = total_weight_of_queen, 
                     color = as.factor(no_of_primary_queen), 
                     fill = as.factor(no_of_primary_queen))) +
  stat_function(fun = function(x){1/(1+exp(-(summary(model1)[[10]][1] + summary(model1)[[10]][2]*x)))}, color = "black") +
  geom_point(size = 2, stroke = 1, shape=21) +
  scale_color_manual(values = c("black", "black")) +
  scale_fill_manual(values = c("white", "black")) +
  scale_y_continuous(breaks=seq(0,1,by=0.25),limits=c(0,1.1),expand=c(0,0)) +
  scale_x_continuous(breaks=seq(0,3000,by=200),limits=c(0,1100),expand=c(0,0)) +
  theme_classic() +
  theme(legend.justification = c(1, 1),　
        legend.text = element_text(size=8), 
        legend.title = element_text(size=8), 
        axis.title.x = element_text(size=8), 
        axis.title.y = element_text(vjust = 0.5,angle=90,size=8), 
        axis.text.x = element_text(size=8,color="black"), 
        axis.text.y = element_text(size=8,color="black"),
        axis.ticks.x = element_line(size = 0.5), 
        axis.line = element_line(size = 0.5, 
                                linetype = "solid", 
                                lineend = "round") 
        ) +
  labs(y = "proportion of the most dominant clone group", 
       x = "total weight of queens (mg)",
       title ="")
Fig_2c

# save figure
# ggsave(file = "Figure 2c.pdf", plot = Fig_2c, dpi = 300, width = 140, height = 89, units = c("mm"))
```


# Comparizon of allele frequencies between driven and non-dvien colonies
## * data loading
```{r}
df <- read_excel("Dataset S1.xlsx", sheet = "alleles in non-driven colonies")
columns_to_process <- c("Rf24_2", "Rf6_1", "Rf21_1", "Rs15", "Rs10", "Rs78", "Rs68", "Rs02")
df_expanded <- df %>%
  separate_rows(all_of(columns_to_process), sep = "/")  %>%
  mutate(across(all_of(columns_to_process), ~as.numeric(.)))

freq <- read_excel("Dataset S1.xlsx", sheet = "alleles in dominated clones") %>% 
  rbind(.,df_expanded) %>% 
  mutate(dominance_level = ifelse(drive_level == "driven", 1, 0))
```

## * Fisher's exact test
### * Rf24_2
```{r}
Rf24_2 <- freq %>%
  dplyr::select(colony, drive_level, Rf24_2)
Rf24_2_dominant <- Rf24_2 %>% 
  filter(drive_level == "driven") %>% 
  group_by(Rf24_2) %>% 
  summarize(number = n()) %>% 
  mutate(drived_allele = dense_rank(-number)) %>% 
  filter(drived_allele == "1") %>% 
  dplyr::select(Rf24_2, drived_allele)

Rf24_2 <- Rf24_2 %>% 
  left_join(Rf24_2_dominant) %>% 
  replace_na(list(drived_allele = 0))

cross_table_Rf24_2 <- table(Rf24_2$drive_level, Rf24_2$drived_allele)

fisher.test(cross_table_Rf24_2)
```

### * Rf6_1
```{r}
Rf6_1 <- freq %>%
  dplyr::select(colony, drive_level, Rf6_1)
Rf6_1_dominant <- Rf6_1 %>% 
  filter(drive_level == "driven") %>% 
  group_by(Rf6_1) %>% 
  summarize(number = n()) %>% 
  mutate(drived_allele = dense_rank(-number)) %>% 
  filter(drived_allele == "1") %>% 
  dplyr::select(Rf6_1, drived_allele)

Rf6_1 <- Rf6_1 %>% 
  left_join(Rf6_1_dominant) %>% 
  replace_na(list(drived_allele = 0))

cross_table_Rf6_1 <- table(Rf6_1$drive_level, Rf6_1$drived_allele)

fisher.test(cross_table_Rf6_1)
```

### * Rf21_1
```{r}
Rf21_1 <- freq %>%
  dplyr::select(colony, drive_level, Rf21_1)
Rf21_1_dominant <- Rf21_1 %>% 
  filter(drive_level == "driven") %>% 
  group_by(Rf21_1) %>% 
  summarize(number = n()) %>% 
  mutate(drived_allele = dense_rank(-number)) %>% 
  filter(drived_allele == "1") %>% 
  dplyr::select(Rf21_1, drived_allele)

Rf21_1 <- Rf21_1 %>% 
  left_join(Rf21_1_dominant) %>% 
  replace_na(list(drived_allele = 0))

cross_table_Rf21_1 <- table(Rf21_1$drive_level, Rf21_1$drived_allele)

fisher.test(cross_table_Rf21_1)
```

### * Rs15
```{r}
Rs15 <- freq %>%
  dplyr::select(colony, drive_level, Rs15)
Rs15_dominant <- Rs15 %>% 
  filter(drive_level == "driven") %>% 
  group_by(Rs15) %>% 
  summarize(number = n()) %>% 
  mutate(drived_allele = dense_rank(-number)) %>% 
  filter(drived_allele == "1") %>% 
  dplyr::select(Rs15, drived_allele)

Rs15 <- Rs15 %>% 
  left_join(Rs15_dominant) %>% 
  replace_na(list(drived_allele = 0))

cross_table_Rs15 <- table(Rs15$drive_level, Rs15$drived_allele)

fisher.test(cross_table_Rs15)
```

### * Rs10
```{r}
Rs10 <- freq %>%
  dplyr::select(colony, drive_level, Rs10)
Rs10_dominant <- Rs10 %>% 
  filter(drive_level == "driven") %>% 
  group_by(Rs10) %>% 
  summarize(number = n()) %>% 
  mutate(drived_allele = dense_rank(-number)) %>% 
  filter(drived_allele == "1") %>% 
  dplyr::select(Rs10, drived_allele)

Rs10 <- Rs10 %>% 
  left_join(Rs10_dominant) %>% 
  replace_na(list(drived_allele = 0))

cross_table_Rs10 <- table(Rs10$drive_level, Rs10$drived_allele)

fisher.test(cross_table_Rs10)
```

### * Rs78
```{r}
Rs78 <- freq %>%
  dplyr::select(colony, drive_level, Rs78)
Rs78_dominant <- Rs78 %>% 
  filter(drive_level == "driven") %>% 
  group_by(Rs78) %>% 
  summarize(number = n()) %>% 
  mutate(drived_allele = dense_rank(-number)) %>% 
  filter(drived_allele == "1") %>% 
  dplyr::select(Rs78, drived_allele)

Rs78 <- Rs78 %>% 
  left_join(Rs78_dominant) %>% 
  replace_na(list(drived_allele = 0))

cross_table_Rs78 <- table(Rs78$drive_level, Rs78$drived_allele)

fisher.test(cross_table_Rs78)
```

### * Rs68
```{r}
Rs68 <- freq %>%
  dplyr::select(colony, drive_level, Rs68)
Rs68_dominant <- Rs68 %>% 
  filter(drive_level == "driven") %>% 
  group_by(Rs68) %>% 
  summarize(number = n()) %>% 
  mutate(drived_allele = dense_rank(-number)) %>% 
  filter(drived_allele == "1") %>% 
  dplyr::select(Rs68, drived_allele)

Rs68 <- Rs68 %>% 
  left_join(Rs68_dominant) %>% 
  replace_na(list(drived_allele = 0))

cross_table_Rs68 <- table(Rs68$drive_level, Rs68$drived_allele)

fisher.test(cross_table_Rs68)
```

### * Rs02
```{r}
Rs02 <- freq %>%
  dplyr::select(colony, drive_level, Rs02)
Rs02_dominant <- Rs02 %>% 
  filter(drive_level == "driven") %>% 
  group_by(Rs02) %>% 
  summarize(number = n()) %>% 
  mutate(drived_allele = dense_rank(-number)) %>% 
  filter(drived_allele == "1") %>% 
  dplyr::select(Rs02, drived_allele)

Rs02 <- Rs02 %>% 
  left_join(Rs02_dominant) %>% 
  replace_na(list(drived_allele = 0))

cross_table_Rs02 <- table(Rs02$drive_level, Rs02$drived_allele)

fisher.test(cross_table_Rs02)
```

## * figure
### * histgram of alleles at Rs15 in non-driven colonies
```{r}
list_Rs15 <- freq %>% 
  group_by(Rs15) %>% 
  summarize(count = n()) %>% 
  dplyr::select(Rs15)

summary_Rs15 <- freq %>% 
  group_by(Rs15,drive_level) %>% 
  summarize(count = n())

summary_Rs15_non_driven1 <- summary_Rs15 %>% 
  filter(drive_level == "non_driven")

summary_Rs15_non_driven <- list_Rs15 %>% 
  left_join(summary_Rs15_non_driven1) %>% 
  mutate(drive_level = "non_driven") %>% 
  replace_na(list(count = 0)) %>% 
  mutate(Rs15 = as.numeric(Rs15)) %>% 
  mutate(rank = row_number(Rs15)) %>% 
  mutate(Rs15 = paste(rank,Rs15,sep="_"))

Fig_2d_non_driven <- 
  ggplot(summary_Rs15_non_driven, aes(y = count,
                                        x = Rs15
                                        )) +
  geom_bar(stat = "identity", color="black") + 
  scale_y_continuous(breaks=seq(0,100,by=5),limits=c(0,16)) + 
  labs(y = "no. of alleles", 
       x = "allele",
       title ="alleles at Rs15 in non-driven colonies") +
  theme_classic()
Fig_2d_non_driven

# save figure
# ggsave(file = "Figure 2d non-driven.pdf", plot = Fig_2d_non_driven, dpi = 300, width = 140, height = 89, units = c("mm"))
```

### * histgram of alleles at Rs15 in driven colonies
```{r}
summary_Rs15_driven1 <- summary_Rs15 %>% 
  filter(drive_level == "driven")

summary_Rs15_driven <- list_Rs15 %>% 
  left_join(summary_Rs15_driven1) %>% 
  mutate(drive_level = "driven") %>% 
  replace_na(list(count = 0)) %>% 
  mutate(Rs15 = as.numeric(Rs15)) %>% 
  mutate(rank = row_number(Rs15)) %>% 
  mutate(Rs15 = paste(rank,Rs15,sep="_"))

Fig_2d_driven <- 
  ggplot(summary_Rs15_driven, aes(y = count,
                                    x = Rs15
                                    )) +
  geom_bar(stat = "identity", color="black") + 
  scale_y_continuous(breaks=seq(0,100,by=1),limits=c(0,4.2)) + 
  labs(y = "no. of alleles", 
       x = "allele",
       title ="alleles at Rs15 in driven colonies") +
  theme_classic()
Fig_2d_driven

# save figure
# ggsave(file = "Figure 2d driven.pdf", plot = Fig_2d_driven, dpi = 300, width = 140, height = 89, units = c("mm"))
```

### * histgram of alleles at Rs68 in non-driven colonies
```{r}
list_Rs68 <- freq %>% 
  group_by(Rs68) %>% 
  summarize(count = n()) %>% 
  dplyr::select(Rs68)

summary_Rs68 <- freq %>% 
  group_by(Rs68,drive_level) %>% 
  summarize(count = n())

summary_Rs68_non_driven1 <- summary_Rs68 %>% 
  filter(drive_level == "non_driven")

summary_Rs68_non_driven <- list_Rs68 %>% 
  left_join(summary_Rs68_non_driven1) %>% 
  mutate(drive_level = "non_driven") %>% 
  replace_na(list(count = 0)) %>% 
  mutate(Rs68 = as.numeric(Rs68)) %>% 
  mutate(rank = row_number(Rs68)) %>% 
  mutate(Rs68 = paste(rank,Rs68,sep="_"))

Fig_2e_non_driven <- 
  ggplot(summary_Rs68_non_driven, aes(y = count,
                                        x = Rs68
                                        )) +
  geom_bar(stat = "identity", color="black") + 
  scale_y_continuous(breaks=seq(0,100,by=5),limits=c(0,16)) + 
  labs(y = "no. of alleles", 
       x = "allele",
       title ="alleles at Rs68 in non-driven colonies") +
  theme_classic()
Fig_2e_non_driven

# save figure
# ggsave(file = "Figure 2e non-driven.pdf", plot = Fig_2e_non_driven, dpi = 300, width = 140, height = 89, units = c("mm"))
```

### * histgram of alleles at Rs68 in driven colonies
```{r}
summary_Rs68_driven1 <- summary_Rs68 %>% 
  filter(drive_level == "driven")

summary_Rs68_driven <- list_Rs68 %>% 
  left_join(summary_Rs68_driven1) %>% 
  mutate(drive_level = "driven") %>% 
  replace_na(list(count = 0)) %>% 
  mutate(Rs68 = as.numeric(Rs68)) %>% 
  mutate(rank = row_number(Rs68)) %>% 
  mutate(Rs68 = paste(rank,Rs68,sep="_"))

Fig_2e_driven <- 
  ggplot(summary_Rs68_driven, aes(y = count,
                                    x = Rs68
                                    )) +
  geom_bar(stat = "identity", color="black") + 
  scale_y_continuous(breaks=seq(0,100,by=1),limits=c(0,2.2)) + 
  labs(y = "no. of alleles", 
       x = "allele",
       title ="alleles at Rs68 in driven colonies") +
  theme_classic()
Fig_2e_driven

# save figure
# ggsave(file = "Figure 2e driven.pdf", plot = Fig_2e_driven, dpi = 300, width = 140, height = 89, units = c("mm"))
```


# Heterozygosity of males and females of pre-founding population
## * data loading
```{r}
Fig3 <- read_excel("Dataset S1.xlsx", sheet="genotype of alates") %>% 
  as.data.frame(.) 
Fig3a <- Fig3 %>% 
  filter(stage == "pre_founding")
Fig3a_male <- Fig3a %>% 
  filter(sex == "Male") %>% 
  mutate(loci = if_else(heterozygosity <= 0, "homo", "hetero"))
Fig3a_female <- Fig3a %>% 
  filter(sex == "Female")
```

## * sample size
```{r}
summary_Fig3 <- Fig3 %>% 
  group_by(sex, stage, parthenogenesis) %>% 
  summarize(number = n())
summary_Fig3

summary_Fig3a <- Fig3a %>% 
  group_by(sex) %>% 
  summarize(number = n())
summary_Fig3a
```

## * statistical analysis
### * comparison of the frequency of asexual females between pre- and post-foundation population
```{r}
summary_data <- summary_Fig3 %>% 
  filter(sex == "Female") %>% 
  pivot_wider(., names_from = "parthenogenesis", values_from = "number") 
fisher_data <- matrix(c(summary_data$sexual, summary_data$asexual), ncol=2, nrow=2, byrow=F)
fisher.test(fisher_data)
```

### * comparison of head width in males in pre-foundation population
```{r}
model_freqaf1 <- glm(head_width ~ loci, family = gaussian, Fig3a_male)
model_freqaf2 <- glm(head_width ~ 1, family = gaussian, Fig3a_male)
summary(model_freqaf1)
Anova(model_freqaf1)
anova(model_freqaf1, model_freqaf2, test="Chi")
```

## * figure
```{r}
# list
heterozygosity <- c("0","0.125","0.25","0.375","0.5","0.625","0.75","0.875","1")
list <- data.frame(heterozygosity) %>% 
  mutate(heterozygosity = as.factor(heterozygosity))

# male
summary_Fig3a_male_raw <- Fig3a_male %>% 
  group_by(heterozygosity) %>% 
  summarize(number = n()) %>% 
  mutate(heterozygosity = as.factor(heterozygosity))
summary_Fig3a_male <- list %>% 
  left_join(summary_Fig3a_male_raw, by="heterozygosity")
  
Fig_3a_male <- 
  ggplot(summary_Fig3a_male, aes(y = number,
                                 x = as.factor(heterozygosity)
                                 )) +
  geom_bar(stat = "identity", color="black") + 
  scale_y_continuous(breaks=seq(0,100,by=10),limits=c(0,51)) + 
  labs(y = "no. of individuals", 
       x = "heterozygosity",
       title ="males of pre-founding population") +
  theme_classic()
Fig_3a_male


# female
summary_Fig3a_female_raw <- Fig3a_female %>% 
  group_by(heterozygosity) %>% 
  summarize(number = n()) %>% 
  mutate(heterozygosity = as.factor(heterozygosity))
summary_Fig3a_female <- list %>% 
  left_join(summary_Fig3a_female_raw)

Fig_3a_female <- 
  ggplot(summary_Fig3a_female, aes(y = number,
                                   x = as.factor(heterozygosity)
                                   )) +
  geom_bar(stat = "identity", color="black") + 
  scale_y_continuous(breaks=seq(0,100,by=10),limits=c(0,51)) + 
  labs(y = "no. of individuals", 
       x = "heterozygosity",
       title ="females of pre-founding population") +
  theme_classic()
Fig_3a_female

# save data
# ggsave(file = "Figure 3a male.pdf", plot = Fig_3a_male, dpi = 300, width = 140, height = 89, units = c("mm"))
# ggsave(file = "Figure 3a female.pdf", plot = Fig_3a_female, dpi = 300, width = 140, height = 89, units = c("mm"))
```

# Heterozygosity of males and females of post-founding population
## * data loading
```{r}
Fig3b <- read_excel("Dataset S1.xlsx", sheet="genotype of alates") %>% 
  as.data.frame(.) %>% 
  filter(stage == "post_founding")
Fig3b_male <- Fig3b %>% 
  filter(sex == "Male")
Fig3b_female <- Fig3b %>% 
  filter(sex == "Female")
```

## * sample size
```{r}
summary_Fig3b <- Fig3b %>% 
  group_by(sex) %>% 
  summarize(number = n())
summary_Fig3b
```

## * figure
```{r}
# male
summary_Fig3b_male_raw <- Fig3b_male %>% 
  group_by(heterozygosity) %>% 
  summarize(number = n()) %>% 
  mutate(heterozygosity = as.factor(heterozygosity))
summary_Fig3b_male <- list %>% 
  left_join(summary_Fig3b_male_raw, by="heterozygosity")
  
Fig_3b_male <- 
  ggplot(summary_Fig3b_male, aes(y = number,
                                 x = as.factor(heterozygosity)
                                 )) +
  geom_bar(stat = "identity", color="black") + 
  scale_y_continuous(breaks=seq(0,100,by=10),limits=c(0,51)) + 
  labs(y = "no. of individuals", 
       x = "heterozygosity",
       title ="males of pre-founding population") +
  theme_classic()
Fig_3b_male


# female
summary_Fig3b_female_raw <- Fig3b_female %>% 
  group_by(heterozygosity) %>% 
  summarize(number = n()) %>% 
  mutate(heterozygosity = as.factor(heterozygosity))
summary_Fig3b_female <- list %>% 
  left_join(summary_Fig3b_female_raw)

Fig_3b_female <- 
  ggplot(summary_Fig3b_female, aes(y = number,
                                   x = as.factor(heterozygosity)
                                   )) +
  geom_bar(stat = "identity", color="black") + 
  scale_y_continuous(breaks=seq(0,100,by=10),limits=c(0,51)) + 
  labs(y = "no. of individuals", 
       x = "heterozygosity",
       title ="females of pre-founding population") +
  theme_classic()
Fig_3b_female

# save figure
# ggsave(file = "Figure 3b male.pdf", plot = Fig_3b_male, dpi = 300, width = 140, height = 89, units = c("mm"))
# ggsave(file = "Figure 3b female.pdf", plot = Fig_3b_female, dpi = 300, width = 140, height = 89, units = c("mm"))
```

# Comparisons of head widths of pre- and post-foundation individuals for both sexes
## * data loading
```{r}
Fig4ab <- read_excel("Dataset S1.xlsx", sheet="genotype of alates") %>% 
  as.data.frame(.) %>% 
  mutate(category = paste(sex, stage, sep="_"))
Fig4ab_male <- Fig4ab %>% 
  filter(sex == "Male")
Fig4ab_female <- Fig4ab %>% 
  filter(sex == "Female")

Fig4ab_male_pre <- Fig4ab_male %>%
  filter(stage=="pre_founding")
Fig4ab_male_post <- Fig4ab_male %>%
  filter(stage=="post_founding")
Fig4ab_female_pre <- Fig4ab_female %>%
  filter(stage=="pre_founding")
Fig4ab_female_post <- Fig4ab_female %>%
  filter(stage=="post_founding")
```

## * sample size
```{r}
summary_Fig4a <- Fig4ab %>% 
  group_by(category) %>% 
  summarize(number = n())
summary_Fig4a
```

## * statistical analysis
### * detection of sex ratio bias
```{r}
# whole data
chisq.test(c(43,111), p=c(1/2,1/2))

# after removing parthenogenetically-produced females
chisq.test(c(43,72.7), p=c(1/2,1/2))
```

### * comparison of head width between pre- and post-foundation population in males
```{r}
# Shapiro-Wilk test
shapiro.test(Fig4ab_male_pre$head_width)
shapiro.test(Fig4ab_male_post$head_width)

# t test
t.test(head_width ~ stage, Fig4ab_male)
```

### * comparison of head width between pre- and post-foundation population in females
```{r}
# Shapiro-Wilk test
shapiro.test(Fig4ab_female_pre$head_width)
shapiro.test(Fig4ab_female_post$head_width)

#  Wilcoxon signed-rank sum test
wilcox.test(head_width ~ stage, correct=FALSE, Fig4ab_female)
```

## * figure
```{r}
Fig4ab$category <- factor(Fig4ab$category, levels=c("Male_pre_founding","Male_post_founding","Female_pre_founding","Female_post_founding"))

Fig_4a <-
ggplot(Fig4ab, aes(y = head_width, x = factor(category))) + 
  geom_boxplot(width=0.6) + 
  #scale_y_continuous(breaks=seq(0,2,by=0.05),limits=c(0.85,1.05)) +
  labs(y = "head width (mm)", 
       x = "",
       title ="") +
  theme_classic()
Fig_4a

# save figure
# ggsave(file = "Figure 4a.pdf", plot = Fig_4a, dpi = 300, width = 140, height = 89, units = c("mm"))
```

# Comparisons of head widths between heterozygous and totally homozygous males in pre-founding population
## * data loading
```{r}
Fig4b_male <- Fig4ab_male_pre %>% 
  mutate(het = if_else(heterozygosity > 0, "heterozygous", "homozygous"))

Fig4b_male_hetero <- Fig4b_male %>% 
  filter(het == "heterozygous")
Fig4b_male_homo <- Fig4b_male %>% 
  filter(het == "homozygous")
```

## * sample size
```{r}
summary_Fig4b_male <- Fig4b_male %>% 
  group_by(het) %>% 
  summarize(number = n())
summary_Fig4b_male
```

## * statistical analysis
```{r}
# Shapiro-Wilk test
shapiro.test(Fig4b_male_hetero$head_width)
shapiro.test(Fig4b_male_homo$head_width)

#  t-test
t.test(head_width ~ het, Fig4b_male)
```

# Comparisons of head widths between sexual and asexual females in pre-founding population
## * data loading
```{r}
Fig4b <- Fig4ab_female %>% 
  mutate(het = if_else(heterozygosity > 0, "sexual", "asexual")) %>% 
  filter(stage == "pre_founding")

Fig4b_s <- Fig4b %>% 
  filter(het == "sexual")
Fig4b_a <- Fig4b %>% 
  filter(het == "asexual")
```

## * sample size
```{r}
summary_Fig4b <- Fig4b %>% 
  group_by(het) %>% 
  summarize(number = n())
summary_Fig4b
```

## * statistical analysis
```{r}
# Shapiro-Wilk test
shapiro.test(Fig4b_s$head_width)
shapiro.test(Fig4b_a$head_width)

#  Wilcoxon signed-rank sum test
wilcox.test(head_width ~ het, correct=FALSE, Fig4b)
```

## * figure
```{r}
Fig4b$het <- factor(Fig4b$het, levels=c("sexual","asexual"))

Fig_4b <-
ggplot(Fig4b, aes(y = head_width, x = factor(het))) + 
  geom_boxplot(#outlier.shape = NA, 
               width=0.6 
               ) + 
  scale_y_continuous(breaks=seq(0,2,by=0.05),limits=c(0.79,1.06)) +
  labs(y = "head width (mm)", 
       x = "",
       title ="") +
  theme_classic()
Fig_4b

# save figure
# ggsave(file = "Figure 4b.pdf", plot = Fig_4b, dpi = 300, width = 140, height = 89, units = c("mm"))
```

# Comparison of body weights between sexual and asexual alates collected from three natal colonies
## * data loading
```{r}
Fig4c <- read_excel("Dataset S1.xlsx", sheet="survivership of alates") %>% 
  as.data.frame(.) %>% 
  mutate(category = paste(colony, parthenogenesis, sep="_"))

Fig4c$category <- factor(Fig4c$category, levels=c("A_sexual","A_asexual","B_sexual","B_asexual","C_sexual","C_asexual"))

Fig4c_female <- Fig4c %>% 
  filter(sex == "Female")
Fig4c_female_ABC <- Fig4c_female %>% 
  filter(str_detect(colony, "A|B|C"))
Fig4c_female_A <- Fig4c_female %>% 
  filter(colony == "A")
Fig4c_female_A_s <- Fig4c_female_A %>% 
  filter(parthenogenesis == "sexual")
Fig4c_female_A_a <- Fig4c_female_A %>% 
  filter(parthenogenesis == "asexual")
Fig4c_female_B <- Fig4c_female %>% 
  filter(colony == "B")
Fig4c_female_B_s <- Fig4c_female_B %>% 
  filter(parthenogenesis == "sexual")
Fig4c_female_B_a <- Fig4c_female_B %>% 
  filter(parthenogenesis == "asexual")
Fig4c_female_C <- Fig4c_female %>% 
  filter(colony == "C")
Fig4c_female_C_s <- Fig4c_female_C %>% 
  filter(parthenogenesis == "sexual")
Fig4c_female_C_a <- Fig4c_female_C %>% 
  filter(parthenogenesis == "asexual")
```

## * sample size
```{r}
summary_Fig4c <- Fig4c %>% 
  group_by(colony, sex, parthenogenesis) %>% 
  summarize(number = n())
summary_Fig4c
```

## * statistical analysis
### * comparison of body weights between sexual and asexual alates in colony A
```{r}
# Shapiro-Wilk test
shapiro.test(Fig4c_female_A_a$weight)
shapiro.test(Fig4c_female_A_s$weight)

# t test
t.test(weight ~ parthenogenesis, Fig4c_female_A)
```

### * comparison of body weights between sexual and asexual alates in colony B
```{r}
# Shapiro-Wilk test
shapiro.test(Fig4c_female_B_a$weight)
shapiro.test(Fig4c_female_B_s$weight)

#  t-test
t.test(weight ~ parthenogenesis, Fig4c_female_B)
```

### * comparison of body weights between sexual and asexual alates in colony C
```{r}
# Shapiro-Wilk test
shapiro.test(Fig4c_female_C_a$weight)
shapiro.test(Fig4c_female_C_s$weight)

#  t-test
t.test(weight ~ parthenogenesis, Fig4c_female_C)
```

## * figure
```{r}
Fig_4c <-
ggplot(Fig4c_female_ABC, aes(y = weight, x = factor(category))) + 
  geom_boxplot(width=0.6) + 
  scale_y_continuous(breaks=seq(0,10,by=0.2),limits=c(2.1,3.7)) +
  labs(y = "freash weight of alates (mg)", 
       x = "",
       title ="") +
  theme_classic()
Fig_4c

# save figure
# ggsave(file = "Figure 4c.pdf", plot = Fig_4c, dpi = 300, width = 140, height = 89, units = c("mm"))
```

# Comparison of survival of males, sexual females, and asexual females
```{r}
# data loading
Fig4d <- read_excel("Dataset S1.xlsx", sheet="survivership of alates") %>% 
  as.data.frame(.)
Fig4d_male <- Fig4d %>% 
  filter(sex == "Male") %>%
  mutate(category = sex)
Fig4d_female <- Fig4d %>% 
  filter(sex == "Female") %>% 
  mutate(category = paste(sex, parthenogenesis, sep="_"))
Fig4d <- rbind(Fig4d_male, Fig4d_female) %>% 
  filter(treatment == "control")
Fig4d_sexual_asexual <- Fig4d %>% 
  filter(!category == "Male")
Fig4d_male_sexual <- Fig4d %>% 
  filter(!category == "Female_asexual")
Fig4d_male_asexual <- Fig4d %>% 
  filter(!category == "Female_sexual")
Fig4d_c_asexual <- Fig4d %>% 
  filter(category == "Female_asexual")
Fig4d_c_sexual <- Fig4d %>% 
  filter(category == "Female_sexual")
Fig4d_c_male <- Fig4d %>% 
  filter(sex == "Male")

Fig4d_t <- rbind(Fig4d_male, Fig4d_female) %>% 
  filter(treatment == "10^5")
Fig4d_t_sexual_asexual <- Fig4d_t %>% 
  filter(!category == "Male")
Fig4d_t_male_sexual <- Fig4d_t %>% 
  filter(!category == "Female_asexual")
Fig4d_t_male_asexual <- Fig4d_t %>% 
  filter(!category == "Female_sexual")
Fig4d_t_sexual <- Fig4d_t %>% 
  filter(category == "Female_sexual")
Fig4d_t_asexual <- Fig4d_t %>% 
  filter(category == "Female_asexual")
Fig4d_t_male <- Fig4d_t %>% 
  filter(category == "Male")
```

## * sample size
```{r}
summary_Fig4d <- Fig4d %>% 
  group_by(category) %>% 
  summarize(number = n())
summary_Fig4d
```

## * statistical analysis
### * comparison of survival in control group
```{r}
survdiff(Surv(event_time, status) ~ category, data = Fig4d_sexual_asexual)
survdiff(Surv(event_time, status) ~ category, data = Fig4d_male_sexual)
survdiff(Surv(event_time, status) ~ category, data = Fig4d_male_asexual)
```

### * bootstrap analysis
```{r}
if("lme4" %in% loadedNamespaces()) detach("package:lme4", unload=TRUE)
if (!requireNamespace("survival", quietly = TRUE)) install.packages("survival")
if (!requireNamespace("boot", quietly = TRUE)) install.packages("boot")

library(survival)
library(boot)

surv_est <- function(data, indices) {
  d <- data[indices, ]
  fit <- survfit(Surv(event_time, status) ~ 1, data = d)
  if (is.na(summary(fit)$table['median'])) { 
    return(max(d$event_time)) 
  } else {
    return(summary(fit)$table['median']) 
  }
}

# control treatment
# asexual female
boot_results <- boot(data = Fig4d_c_asexual, statistic = surv_est, R = 1000)

if(!any(is.na(boot_results$t))) {
  # percentile
  boot_ci_perc <- boot.ci(boot_results, type = "perc")
  print(boot_ci_perc)
  
  # normal
  boot_ci_norm <- boot.ci(boot_results, type = "norm")
  print(boot_ci_norm)
}

# sexual female
boot_results <- boot(data = Fig4d_c_sexual, statistic = surv_est, R = 1000)

if(!any(is.na(boot_results$t))) {
  # percentile
  boot_ci_perc <- boot.ci(boot_results, type = "perc")
  print(boot_ci_perc)
  
  # normal
  boot_ci_norm <- boot.ci(boot_results, type = "norm")
  print(boot_ci_norm)
}

# male
boot_results <- boot(data = Fig4d_c_male, statistic = surv_est, R = 1000)

if(!any(is.na(boot_results$t))) {
  # percentile
  boot_ci_perc <- boot.ci(boot_results, type = "perc")
  print(boot_ci_perc)
  
  # normal
  boot_ci_norm <- boot.ci(boot_results, type = "norm")
  print(boot_ci_norm)
}


# pathogen treatment
boot_results <- boot(data = Fig4d_t_asexual, statistic = surv_est, R = 1000)

if(!any(is.na(boot_results$t))) {
  # percentile
  boot_ci_perc <- boot.ci(boot_results, type = "perc")
  print(boot_ci_perc)
  
  # normal
  boot_ci_norm <- boot.ci(boot_results, type = "norm")
  print(boot_ci_norm)
}

# sexual female
boot_results <- boot(data = Fig4d_t_sexual, statistic = surv_est, R = 1000)

if(!any(is.na(boot_results$t))) {
  # percentile
  boot_ci_perc <- boot.ci(boot_results, type = "perc")
  print(boot_ci_perc)
  
  # normal
  boot_ci_norm <- boot.ci(boot_results, type = "norm")
  print(boot_ci_norm)
}

# male
boot_results <- boot(data = Fig4d_t_male, statistic = surv_est, R = 1000)

if(!any(is.na(boot_results$t))) {
  # percentile
  boot_ci_perc <- boot.ci(boot_results, type = "perc")
  print(boot_ci_perc)
  
  # normal
  boot_ci_norm <- boot.ci(boot_results, type = "norm")
  print(boot_ci_norm)
}
```

### * comparison of survival in pathogen treatment group
```{r}
survdiff(Surv(event_time, status) ~ category, data = Fig4d_t_sexual_asexual)
survdiff(Surv(event_time, status) ~ category, data = Fig4d_t_male_sexual)
survdiff(Surv(event_time, status) ~ category, data = Fig4d_t_male_asexual)
```

## * figure
```{r}
km_fit_sa <- survfit(Surv(event_time, status) ~ category, data = Fig4d_sexual_asexual)
summary(km_fit_sa)

palette <- c("gray", "red")

Figure_Sx2 <- 
  ggsurvplot(km_fit_sa,
             data = Fig4d_sexual_asexual,
             xlab = "Time",
             ylab = "Survival probability",
             break.time.by = 5,
             pval = TRUE,
             conf.int = TRUE,
             palette = palette) # 正しいパレットを使用
Figure_Sx2

# save figure
# ggsave(file = "Figure_Sx2.pdf", plot = Figure_Sx2$plot, dpi = 300, width = 140, height = 89, units = "mm")


# km_fit <- survfit(Surv(event_time, status) ~ category, data = Fig4d)
# 
# summary(km_fit)
# 
# palette <- c("gray", "red", "blue")
# 
# Figure_Sx <- 
#   ggsurvplot(km_fit,
#              data = Fig4d,
#              xlab = "Time",
#              ylab = "Survival probability",
#              break.time.by = 5,
#              pval = TRUE,
#              conf.int = TRUE,
#              palette = palette) # 正しいパレットを使用
# Figure_Sx
# 
# ggsave(file = "Figure_Sx1.pdf", plot = Figure_Sx$plot, dpi = 300, width = 140, height = 89, units = "mm")
```

# Number of young and old secondary queens
```{r}
# data loading
FigS1b <- read_excel("Dataset S1.xlsx", sheet="weight of young and old SQs") %>% 
  as.data.frame(.) %>% 
  filter(caste == "SQ")
SQ_190521I <- FigS1b %>% 
  filter(colony == "190521I")
SQ_180605L <- FigS1b %>% 
  filter(colony == "180605L")
```

## * sample size
```{r}
summary_FigS1b <- FigS1b %>% 
  group_by(colony) %>% 
  summarize(number_of_SQ = n())
summary_FigS1b
```

## * figure
### * histgram of body weight of queens in 180605L
```{r}
width <- 0.5

SQ_hist_180605L <-
  ggplot(SQ_180605L, aes(x = body_weight)) + 
  geom_histogram(boundary = 0, binwidth = width, closed ='left' , fill="black", color="white") +
  scale_x_continuous(expand = c(0,0), breaks = seq(0, 100, by = 5), lim = c(0,31)) +
  scale_y_continuous(expand = c(0,0), 
                     breaks = seq(0, 100, by = 10), lim = c(0,35)
                     ) +
  labs(x = "SQ weight (mg)",
       y = "Count", 
       title ="") + 
  theme_classic(base_family = "Helvetica")
SQ_hist_180605L

# save figure
# ggsave(file = "Figure S1b_1.pdf", plot = SQ_hist_180605L, dpi = 300, width = 140, height = 89, units = c("mm"))
```

### * histgram of body weight of queens in 190521I
```{r}
SQ_hist_190521I <-
  ggplot(SQ_190521I, aes(x = body_weight)) + 
  geom_histogram(boundary = 0, binwidth = width, closed ='left' , fill="black", color="white") +
  scale_x_continuous(expand = c(0,0), breaks = seq(0, 100, by = 5), lim = c(0,16)) +
  scale_y_continuous(expand = c(0,0), 
                     breaks = seq(0, 100, by = 10), lim = c(0,50)
                     ) +
  labs(x = "SQ weight (mg)",
       y = "Count", 
       title ="") + 
  theme_classic(base_family = "Helvetica")
SQ_hist_190521I

# save figure
# ggsave(file = "Figure S1b_2.pdf", plot = SQ_hist_190521I, dpi = 300, width = 140, height = 89, units = c("mm"))
```


# Proportion of male and female dealates in pre-foundation population
```{r}
FigS3a <- read_excel("Dataset S1.xlsx", sheet="genotype of alates") %>% 
  as.data.frame(.) %>% 
  filter(stage == "pre_founding")
```

## * sample size
```{r}
summary_FigS3a <- FigS3a %>% 
  group_by(sex) %>% 
  summarize(number = n())
summary_FigS3a
```

## * figure
```{r}
Fig_S3a <-
ggplot(summary_FigS3a, aes(x="", y=number, fill=sex)) +
  geom_bar(width = 1, stat = "identity") + 
  coord_polar("y", start=0)
Fig_S3a

# save figure
# ggsave(file = "Figure S3a.pdf", plot = Fig_S3a, dpi = 300, width = 140, height = 89, units = c("mm"))
```

# Composition of founding units
```{r}
# data loading
FigS3b <- read_excel("Dataset S1.xlsx", sheet="composition of founding units") %>% 
  as.data.frame(.)
```

## * sample size
```{r}
summary_FigS3b <- FigS3b %>% 
  summarize(sum = sum(number_of_colony))
summary_FigS3b
```

## * figure
```{r}
FigS3b$combination <- factor(FigS3b$combination, levels=c("MMFF","MFFF","MMF","MFF","MM","MF","F","M"))

Fig_S3b <-
ggplot(FigS3b, aes(x="", y=number_of_colony, fill=combination)) +
  geom_bar(width = 1, stat = "identity") + 
  coord_polar("y", start=0)
Fig_S3b

# save figure
# ggsave(file = "Figure S3b.pdf", plot = Fig_S3b, dpi = 300, width = 140, height = 89, units = c("mm"))
```
