---
title: 'Inter-clonal competition over queen succession imposes a cost of parthenogenesis on termite colonies'
output:
  html_document: default
  pdf_document: default
---

# packages
```{r}
install.packages("readxl")
install.packages("ggplot2")
install.packages("tidyverse")
install.packages("dplyr")
install.packages("lme4")
install.packages("car")

library(readxl)
library(ggplot2)
library(tidyverse)
library(lme4)
library(car)
library(survival)
library(scales)
library(survminer)
```

# Distribution of the number of secondary queens per colony
```{r}
# data loading
Fig2a <- read_excel("Dataset.xlsx", sheet="no. of royals") %>% 
  as.data.frame(.)
```

## * sample size
```{r}
summary_Fig2a <- Fig2a %>% 
  group_by(no_of_primary_queen) %>% 
  summarize(number = n())
summary_Fig2a
```

## * statistical analysis
```{r}
model1 <- glm(no_of_secondary_queen ~ as.factor(no_of_primary_queen), family = poisson, Fig2a)
model2 <- glm(no_of_secondary_queen ~ 1, family = poisson, Fig2a)
summary(model1)
Anova(model1)
anova(model1, model2, test="Chi")
```

## * figure
```{r}
xmin <- 0 
xmax <- 320 
xscale <- 50 
binwidth <- 10 
ymin <- 0 
ymax <- 33 
yscale <- 10 

par(ps=18)

Fig_2a <- 
  ggplot(Fig2a, aes(x = no_of_secondary_queen, fill = as.factor(no_of_primary_queen))) +
  geom_histogram(boundary = 0, 
                 binwidth = binwidth, 
                 closed ='left' , 
                 alpha = 1, 
                 color = "black", 
                 position = "identity" 
                 ) +
  scale_fill_manual(values = c("white", "black")) + 
  scale_x_continuous(breaks = seq(xmin, xmax, by = xscale), lim = c(xmin, xmax), expand = c(0,0)) + 
  scale_y_continuous(breaks = seq(ymin, ymax, by = yscale), lim = c(ymin, ymax), expand = c(0,0)) + 
  labs(y = "no. of colonies", 
       x = "no. of secondary queens / colony",
       title ="") +
  theme_classic()
Fig_2a

# save figure
# ggsave(file = "Figure 2a.pdf", plot = Fig_2a, dpi = 300, width = 140, height = 89, units = c("mm"))
```

# Relationship between total weight of secondary queens in a colony and the number of clone types
```{r}
# data loading
Fig2bc <- read_excel("Dataset.xlsx", sheet="no. of clones in SQ population") %>% 
  as.data.frame(.)
```

## * sample size
```{r}
summary_Fig2bc <- Fig2bc %>% 
  group_by(no_of_primary_queen) %>% 
  summarize(number = n())
summary_Fig2bc
```

## * statistical analysis
### * comparison of number of clones between colonies with and without PQ
```{r}
model_cdq1 <- glmer(cbind(no_of_clone_types,no_secondary_queens_genotyped-no_of_clone_types) ~ as.factor(no_of_primary_queen) + (1|colony), family = binomial, Fig2bc)
model_cdq2 <- glmer(cbind(no_of_clone_types,no_secondary_queens_genotyped-no_of_clone_types) ~ 1 + (1|colony), family = binomial, Fig2bc)
summary(model_cdq1)
Anova(model_cdq1)
anova(model_cdq1, model_cdq2, test="Chi")
```

### * relationship between colony size and number of clones in SQ population
```{r}
model_cd1 <- glmer(cbind(no_of_clone_types,no_secondary_queens_genotyped-no_of_clone_types) ~ total_weight_of_queen + (1|colony), family = binomial, Fig2bc)
model_cd2 <- glmer(cbind(no_of_clone_types,no_secondary_queens_genotyped-no_of_clone_types) ~ 1 + (1|colony), family = binomial, Fig2bc)
summary(model_cd1)
Anova(model_cd1)
anova(model_cd1, model_cd2, test="Chi")
```

## * figure
```{r}
Fig_2b <- 
  ggplot(Fig2bc, aes(y = no_of_clone_types, 
                     x = total_weight_of_queen, 
                     fill = as.factor(no_of_primary_queen))) +
  stat_function(fun = function(x){(1/(1+exp(-(summary(model_cd1)[[10]][1] + summary(model_cd1)[[10]][2]*x))))*20}, color = "black") +
  geom_point(size = 2, stroke = 1, shape=21) +
  scale_fill_manual(values = c("white", "black")) +
  scale_y_continuous(breaks=seq(0,20,by=5),limits=c(0,20),expand=c(0,0)) +
  scale_x_continuous(breaks=seq(0,3000,by=200),limits=c(0,1100),expand=c(0,0)) +
  theme_classic() +
  theme(legend.justification = c(1, 1),　
        legend.text = element_text(size=8), 
        legend.title = element_text(size=8), 
        axis.title.x = element_text(size=8), 
        axis.title.y = element_text(vjust = 0.5,angle=90,size=8), 
        axis.text.x = element_text(size=8,color="black"), 
        axis.text.y = element_text(size=8,color="black"),
        axis.ticks.x = element_line(size = 0.5), 
        axis.line = element_line(size = 0.5, 
                                linetype = "solid", 
                                lineend = "round")
        ) +
  labs(y = "no. of clone groups", 
       x = "total weight of queens (mg)",
       title ="")
Fig_2b

# save data
# ggsave(file = "Figure 2b.pdf", plot = Fig_2b, dpi = 300, width = 140, height = 89, units = c("mm"))
```

# Relationship between colony size and the proportion of the most dominant clone
## * statistical analysis
### * comparison of the proportion of the most dominant clone between colonies with and without PQ
```{r}
model_cdq1 <- glmer(cbind(no_of_individuals_of_the_most_dominate_clone_type,no_secondary_queens_genotyped-no_of_individuals_of_the_most_dominate_clone_type) ~ total_weight_of_queen + (1|colony), family = binomial, Fig2bc)
model_cdq2 <- glmer(cbind(no_of_individuals_of_the_most_dominate_clone_type,no_secondary_queens_genotyped-no_of_individuals_of_the_most_dominate_clone_type) ~ 1 + (1|colony), family = binomial, Fig2bc)
summary(model_cdq1)
Anova(model_cdq1)
anova(model_cdq1, model_cdq2, test="Chi")
```

### * relationship between colony size and the proportion of the most dominant clone in SQ population
```{r}
model_cd1 <- glmer(cbind(no_of_individuals_of_the_most_dominate_clone_type,no_secondary_queens_genotyped-no_of_individuals_of_the_most_dominate_clone_type) ~ total_weight_of_queen + (1|colony), family = binomial, Fig2bc)
model_cd2 <- glmer(cbind(no_of_individuals_of_the_most_dominate_clone_type,no_secondary_queens_genotyped-no_of_individuals_of_the_most_dominate_clone_type) ~ 1 + (1|colony), family = binomial, Fig2bc)
summary(model_cd1)
Anova(model_cd1)
anova(model_cd1, model_cd2, test="Chi")
```

## * figure
```{r}
Fig_2c <- 
  ggplot(Fig2bc, aes(y = percent_of_the_most_domonate_clone_type/100, 
                     x = total_weight_of_queen, 
                     color = as.factor(no_of_primary_queen), 
                     fill = as.factor(no_of_primary_queen))) +
  stat_function(fun = function(x){1/(1+exp(-(summary(model1)[[10]][1] + summary(model1)[[10]][2]*x)))}, color = "black") +
  geom_point(size = 2, stroke = 1, shape=21) +
  scale_color_manual(values = c("black", "black")) +
  scale_fill_manual(values = c("white", "black")) +
  scale_y_continuous(breaks=seq(0,1,by=0.25),limits=c(0,1.1),expand=c(0,0)) +
  scale_x_continuous(breaks=seq(0,3000,by=200),limits=c(0,1100),expand=c(0,0)) +
  theme_classic() +
  theme(#legend.position = c(1, 1), #レジェンド(横についてるやつ)を上詰め(1)、右詰め(1)にする "bottom"
        legend.justification = c(1, 1),　#レジェンドをちょっと内側に
        legend.text = element_text(size=8), #レジェンドの内容の文字サイズを変更
        legend.title = element_text(size=8), #レジェンドのタイトルサイズを変更
        axis.title.x = element_text(size=8), #x軸タイトルを上詰め(0)、黒色、サイズ１４で
        axis.title.y = element_text(vjust = 0.5,angle=90,size=8), #vjustで上下の位置調整、英語の場合はangle = 90で
        axis.text.x = element_text(size=8,color="black"), #x軸の目盛りに対応した文字サイズを12で
        axis.text.y = element_text(size=8,color="black"),
        axis.ticks.x = element_line(size = 0.5), #x軸の目盛り線をゼロにして消す
        axis.line = element_line(size = 0.5, #軸の線を0.5の太さで
                                linetype = "solid", #線のタイプは実線
                                lineend = "round") #軸の線の終わりは整数値で丸める
        ) +
  labs(y = "proportion of the most dominant clone group", 
       x = "total weight of queens (mg)",
       title ="")
Fig_2c

# save figure
# ggsave(file = "Figure 2c.pdf", plot = Fig_2c, dpi = 300, width = 140, height = 89, units = c("mm"))
```

# Heterozygosity of Males and females of pre-founding population
```{r}
# loading data
Fig3 <- read_excel("Dataset.xlsx", sheet="genotype of alates") %>% 
  as.data.frame(.) 
Fig3a <- Fig3 %>% 
  filter(stage == "pre_founding")
Fig3a_male <- Fig3a %>% 
  filter(sex == "Male") %>% 
  mutate(loci = if_else(heterozygosity <= 0, "homo", "hetero"))
Fig3a_female <- Fig3a %>% 
  filter(sex == "Female")
```

## * sample size
```{r}
summary_Fig3 <- Fig3 %>% 
  group_by(sex, stage, parthenogenesis) %>% 
  summarize(number = n())
summary_Fig3

summary_Fig3a <- Fig3a %>% 
  group_by(sex) %>% 
  summarize(number = n())
summary_Fig3a
```

## * statistical analysis
### * comparison of the frequency of asexual females between pre- and post-foundation population
```{r}
summary_data <- summary_Fig3 %>% 
  filter(sex == "Female") %>% 
  pivot_wider(., names_from = "parthenogenesis", values_from = "number") 
fisher_data <- matrix(c(summary_data$sexual, summary_data$asexual), ncol=2, nrow=2, byrow=F)
fisher.test(fisher_data)
```

### * comparison of head width in males in pre-foundation population
```{r}
model_freqaf1 <- glm(head_width ~ loci, family = gaussian, Fig3a_male)
model_freqaf2 <- glm(head_width ~ 1, family = gaussian, Fig3a_male)
summary(model_freqaf1)
Anova(model_freqaf1)
anova(model_freqaf1, model_freqaf2, test="Chi")
```

## * figure
```{r}
# list
heterozygosity <- c("0","0.125","0.25","0.375","0.5","0.625","0.75","0.875","1")
list <- data.frame(heterozygosity) %>% 
  mutate(heterozygosity = as.factor(heterozygosity))

# male
summary_Fig3a_male_raw <- Fig3a_male %>% 
  group_by(heterozygosity) %>% 
  summarize(number = n()) %>% 
  mutate(heterozygosity = as.factor(heterozygosity))
summary_Fig3a_male <- list %>% 
  left_join(summary_Fig3a_male_raw, by="heterozygosity")
  
Fig_3a_male <- 
  ggplot(summary_Fig3a_male, aes(y = number,
                                 x = as.factor(heterozygosity)
                                 )) +
  geom_bar(stat = "identity", color="black") + 
  scale_y_continuous(breaks=seq(0,100,by=10),limits=c(0,50)) + 
  labs(y = "no. of individuals", 
       x = "heterozygosity",
       title ="males of pre-founding population") +
  theme_classic()
Fig_3a_male


# female
summary_Fig3a_female_raw <- Fig3a_female %>% 
  group_by(heterozygosity) %>% 
  summarize(number = n()) %>% 
  mutate(heterozygosity = as.factor(heterozygosity))
summary_Fig3a_female <- list %>% 
  left_join(summary_Fig3a_female_raw)

Fig_3a_female <- 
  ggplot(summary_Fig3a_female, aes(y = number,
                                   x = as.factor(heterozygosity)
                                   )) +
  geom_bar(stat = "identity", color="black") + 
  scale_y_continuous(breaks=seq(0,100,by=10),limits=c(0,50)) + 
  labs(y = "no. of individuals", 
       x = "heterozygosity",
       title ="females of pre-founding population") +
  theme_classic()
Fig_3a_female

# save data
# ggsave(file = "Figure 3a male.pdf", plot = Fig_3a_male, dpi = 300, width = 140, height = 89, units = c("mm"))
# ggsave(file = "Figure 3a female.pdf", plot = Fig_3a_female, dpi = 300, width = 140, height = 89, units = c("mm"))
```

# Heterozygosity of Males and females of post-founding population
```{r}
# data loading
Fig3b <- read_excel("Dataset.xlsx", sheet="genotype of alates") %>% 
  as.data.frame(.) %>% 
  filter(stage == "post_founding")
Fig3b_male <- Fig3b %>% 
  filter(sex == "Male")
Fig3b_female <- Fig3b %>% 
  filter(sex == "Female")
```

## * sample size
```{r}
summary_Fig3b <- Fig3b %>% 
  group_by(sex) %>% 
  summarize(number = n())
summary_Fig3b
```

## * figure
```{r}
# male
summary_Fig3b_male_raw <- Fig3b_male %>% 
  group_by(heterozygosity) %>% 
  summarize(number = n()) %>% 
  mutate(heterozygosity = as.factor(heterozygosity))
summary_Fig3b_male <- list %>% 
  left_join(summary_Fig3b_male_raw, by="heterozygosity")
  
Fig_3b_male <- 
  ggplot(summary_Fig3b_male, aes(y = number,
                                 x = as.factor(heterozygosity)
                                 )) +
  geom_bar(stat = "identity", color="black") + 
  scale_y_continuous(breaks=seq(0,100,by=10),limits=c(0,50)) + 
  labs(y = "no. of individuals", 
       x = "heterozygosity",
       title ="males of pre-founding population") +
  theme_classic()
Fig_3b_male


# female
summary_Fig3b_female_raw <- Fig3b_female %>% 
  group_by(heterozygosity) %>% 
  summarize(number = n()) %>% 
  mutate(heterozygosity = as.factor(heterozygosity))
summary_Fig3b_female <- list %>% 
  left_join(summary_Fig3b_female_raw)

Fig_3b_female <- 
  ggplot(summary_Fig3b_female, aes(y = number,
                                   x = as.factor(heterozygosity)
                                   )) +
  geom_bar(stat = "identity", color="black") + 
  scale_y_continuous(breaks=seq(0,100,by=10),limits=c(0,50)) + 
  labs(y = "no. of individuals", 
       x = "heterozygosity",
       title ="females of pre-founding population") +
  theme_classic()
Fig_3b_female

# save figure
# ggsave(file = "Figure 3b male.pdf", plot = Fig_3b_male, dpi = 300, width = 140, height = 89, units = c("mm"))
# ggsave(file = "Figure 3b female.pdf", plot = Fig_3b_female, dpi = 300, width = 140, height = 89, units = c("mm"))
```

# Comparisons of head widths of pre- and post-foundation individuals for both sexes
```{r}
# data loading
Fig4ab <- read_excel("Dataset.xlsx", sheet="genotype of alates") %>% 
  as.data.frame(.) %>% 
  mutate(category = paste(sex, stage, sep="_"))
Fig4ab_male <- Fig4ab %>% 
  filter(sex == "Male")
Fig4ab_female <- Fig4ab %>% 
  filter(sex == "Female")
```

## * sample size
```{r}
summary_Fig4a <- Fig4ab %>% 
  group_by(category) %>% 
  summarize(number = n())
summary_Fig4a
```

## * statistical analysis
### * detection of sex ratio bias
```{r}
# whole data
chisq.test(c(43,111), p=c(1/2,1/2))

# after removing parthenogenetically-produced females
chisq.test(c(43,72.7), p=c(1/2,1/2))
```

### * comparison of head width between pre- and post-foundation population in males
```{r}
model_hm1 <- glm(head_width ~ stage, family = gaussian, Fig4ab_male)
model_hm2 <- glm(head_width ~ 1, family = gaussian, Fig4ab_male)
summary(model_hm1)
Anova(model_hm1)
anova(model_hm1, model_hm2, test="Chi")
```

### * comparison of head width between pre- and post-foundation population in females
```{r}
model_hf1 <- glm(head_width ~ stage, family = gaussian, Fig4ab_female)
model_hf2 <- glm(head_width ~ 1, family = gaussian, Fig4ab_female)
summary(model_hf1)
Anova(model_hf1)
anova(model_hf1, model_hf2, test="Chi")
```

## * figure
```{r}
Fig4ab$category <- factor(Fig4ab$category, levels=c("Male_pre_founding","Male_post_founding","Female_pre_founding","Female_post_founding"))

Fig_4a <-
ggplot(Fig4ab, aes(y = head_width, x = factor(category))) + 
  geom_boxplot(width=0.6) + 
  #scale_y_continuous(breaks=seq(0,2,by=0.05),limits=c(0.85,1.05)) +
  labs(y = "head width (mm)", 
       x = "",
       title ="") +
  theme_classic()
Fig_4a

# save figure
# ggsave(file = "Figure 4a.pdf", plot = Fig_4a, dpi = 300, width = 140, height = 89, units = c("mm"))
```

# Comparisons of head widths between sexual and asexual females in pre-founding population
```{r}
# data loading
Fig4b <- Fig4ab_female %>% 
  mutate(het = if_else(heterozygosity > 0, "sexual", "asexual")) %>% 
  filter(stage == "pre_founding")
```

## * sample size
```{r}
summary_Fig4b <- Fig4b %>% 
  group_by(het) %>% 
  summarize(number = n())
summary_Fig4b
```

## * statistical analysis
```{r}
model_hsa1 <- glm(head_width ~ het, family = gaussian, Fig4b)
model_hsa2 <- glm(head_width ~ 1, family = gaussian, Fig4b)
summary(model_hsa1)
Anova(model_hsa1)
anova(model_hsa1, model_hsa2, test="Chi")
```

## * figure
```{r}
Fig4b$het <- factor(Fig4b$het, levels=c("sexual","asexual"))

Fig_4b <-
ggplot(Fig4b, aes(y = head_width, x = factor(het))) + 
  geom_boxplot(#outlier.shape = NA, 
               width=0.6 
               ) + 
  scale_y_continuous(breaks=seq(0,2,by=0.05),limits=c(0.79,1.06)) +
  labs(y = "head width (mm)", 
       x = "",
       title ="") +
  theme_classic()
Fig_4b

# save figure
# ggsave(file = "Figure 4b.pdf", plot = Fig_4b, dpi = 300, width = 140, height = 89, units = c("mm"))
```

# Comparison of body weights between sexual and asexual alates collected from three natal colonies
```{r}
# data loading
Fig4c <- read_excel("Dataset.xlsx", sheet="survivership of alates") %>% 
  as.data.frame(.) %>% 
  mutate(category = paste(colony, parthenogenesis, sep="_"))
```

## * sample size
```{r}
summary_Fig4c <- Fig4c %>% 
  group_by(colony, sex, parthenogenesis) %>% 
  summarize(number = n())
summary_Fig4c
```

## * statistical analysis
### * comparison of body weights between sexual and asexual alates (whole data)
```{r}
model_bww1 <- glmer(weight ~ parthenogenesis + (1|colony), family = gaussian, Fig4c_female_ABC)
model_bww2 <- glmer(weight ~ 1 + (1|colony), family = gaussian, Fig4c_female_ABC)
summary(model_bww1)
Anova(model_bww1)
anova(model_bww1, model_bww2, test="Chi")
```

### * comparison of body weights between sexual and asexual alates in colony A
```{r}
model_bwa1 <- glm(weight ~ parthenogenesis, family = gaussian, Fig4c_female_A)
model_bwa2 <- glm(weight ~ 1, family = gaussian, Fig4c_female_A)
summary(model_bwa1)
Anova(model_bwa1)
anova(model_bwa1, model_bwa2, test="Chi")
```

### * comparison of body weights between sexual and asexual alates in colony B
```{r}
model_bwb1 <- glm(weight ~ parthenogenesis, family = gaussian, Fig4c_female_B)
model_bwb2 <- glm(weight ~ 1, family = gaussian, Fig4c_female_B)
summary(model_bwb1)
Anova(model_bwb1)
anova(model_bwb1, model_bwb2, test="Chi")
```

### * comparison of body weights between sexual and asexual alates in colony C
```{r}
model_bwc1 <- glm(weight ~ parthenogenesis, family = gaussian, Fig4c_female_C)
model_bwc2 <- glm(weight ~ 1, family = gaussian, Fig4c_female_C)
summary(model_bwc1)
Anova(model_bwc1)
anova(model_bwc1, model_bwc2, test="Chi")
```

## * figure
```{r}
Fig4c$category <- factor(Fig4c$category, levels=c("A_sexual","A_asexual","B_sexual","B_asexual","C_sexual","C_asexual"))

Fig4c_female <- Fig4c %>% 
  filter(sex == "Female")
Fig4c_female_ABC <- Fig4c_female %>% 
  filter(str_detect(colony, "A|B|C"))
Fig4c_female_A <- Fig4c_female %>% 
  filter(colony == "A")
Fig4c_female_B <- Fig4c_female %>% 
  filter(colony == "B")
Fig4c_female_C <- Fig4c_female %>% 
  filter(colony == "C")

Fig_4c <-
ggplot(Fig4c_female_ABC, aes(y = weight, x = factor(category))) + 
  geom_boxplot(width=0.6) + 
  scale_y_continuous(breaks=seq(0,10,by=0.2),limits=c(2.1,3.7)) +
  labs(y = "freash weight of alates (mg)", 
       x = "",
       title ="") +
  theme_classic()
Fig_4c

# save figure
# ggsave(file = "Figure 4c.pdf", plot = Fig_4c, dpi = 300, width = 140, height = 89, units = c("mm"))
```

# Comparison of survival of males, sexual females, and asexual females
```{r}
# data loading
Fig4d <- read_excel("Dataset.xlsx", sheet="survivership of alates") %>% 
  as.data.frame(.)
Fig4d_male <- Fig4d %>% 
  filter(sex == "Male") %>%
  mutate(category = sex)
Fig4d_female <- Fig4d %>% 
  filter(sex == "Female") %>% 
  mutate(category = paste(sex, parthenogenesis, sep="_"))
Fig4d <- rbind(Fig4d_male, Fig4d_female) %>% 
  filter(treatment == "control")
Fig4d_sexual_asexual <- Fig4d %>% 
  filter(!category == "Male")
Fig4d_male_sexual <- Fig4d %>% 
  filter(!category == "Female_asexual")
Fig4d_male_asexual <- Fig4d %>% 
  filter(!category == "Female_sexual")

Fig4d_t <- rbind(Fig4d_male, Fig4d_female) %>% 
  filter(treatment == "10^5")
Fig4d_t_sexual_asexual <- Fig4d_t %>% 
  filter(!category == "Male")
Fig4d_t_male_sexual <- Fig4d_t %>% 
  filter(!category == "Female_asexual")
Fig4d_t_male_asexual <- Fig4d_t %>% 
  filter(!category == "Female_sexual")
```

## * sample size
```{r}
summary_Fig4d <- Fig4d %>% 
  group_by(category) %>% 
  summarize(number = n())
summary_Fig4d
```

## * statistical analysis
### * comparison of survival in control group
```{r}
survdiff(Surv(event_time, status) ~ category, data = Fig4d_sexual_asexual)
survdiff(Surv(event_time, status) ~ category, data = Fig4d_male_sexual)
survdiff(Surv(event_time, status) ~ category, data = Fig4d_male_asexual)
```

### * comparison of survival in pathogen treatment group
```{r}
survdiff(Surv(event_time, status) ~ category, data = Fig4d_t_sexual_asexual)
survdiff(Surv(event_time, status) ~ category, data = Fig4d_t_male_sexual)
survdiff(Surv(event_time, status) ~ category, data = Fig4d_t_male_asexual)
```

## * figure
```{r}
survivfit <- survfit(Surv(event_time, status) ~ category, data = Fig4d)
Fig_4d <- ggsurvplot(survivfit, xlim = c(0, 30), expand = FALSE)
Fig_4d$plot <- Fig_4d$plot +
  coord_cartesian(xlim = c(0, 30), expand = FALSE) +
  theme(plot.margin = margin(5.5, 20, 5.5, 5.5))
Fig_4d
```

# Number of young and old secondary queens
```{r}
# data loading
FigS1b <- read_excel("Dataset.xlsx", sheet="weight of young and old SQs") %>% 
  as.data.frame(.) %>% 
  filter(caste == "SQ")
SQ_190521I <- FigS1b %>% 
  filter(colony == "190521I")
SQ_180605L <- FigS1b %>% 
  filter(colony == "180605L")
```

## * sample size
```{r}
summary_FigS1b <- FigS1b %>% 
  group_by(colony) %>% 
  summarize(number_of_SQ = n())
summary_FigS1b
```

## * figure
### * histgram of body weight of queens in 180605L
```{r}
width <- 0.5

SQ_hist_180605L <-
  ggplot(SQ_180605L, aes(x = body_weight)) + 
  geom_histogram(boundary = 0, binwidth = width, closed ='left' , fill="black", color="white") +
  scale_x_continuous(expand = c(0,0), breaks = seq(0, 100, by = 5), lim = c(0,31)) +
  scale_y_continuous(expand = c(0,0), 
                     breaks = seq(0, 100, by = 10), lim = c(0,35)
                     ) +
  labs(x = "SQ weight (mg)",
       y = "Count", 
       title ="") + 
  theme_classic(base_family = "Helvetica")
SQ_hist_180605L

# save figure
# ggsave(file = "Figure S1b_1.pdf", plot = SQ_hist_180605L, dpi = 300, width = 140, height = 89, units = c("mm"))
```

### * histgram of body weight of queens in 190521I
```{r}
SQ_hist_190521I <-
  ggplot(SQ_190521I, aes(x = body_weight)) + 
  geom_histogram(boundary = 0, binwidth = width, closed ='left' , fill="black", color="white") +
  scale_x_continuous(expand = c(0,0), breaks = seq(0, 100, by = 5), lim = c(0,16)) +
  scale_y_continuous(expand = c(0,0), 
                     breaks = seq(0, 100, by = 10), lim = c(0,50)
                     ) +
  labs(x = "SQ weight (mg)",
       y = "Count", 
       title ="") + 
  theme_classic(base_family = "Helvetica")
SQ_hist_190521I

# save figure
# ggsave(file = "Figure S1b_2.pdf", plot = SQ_hist_190521I, dpi = 300, width = 140, height = 89, units = c("mm"))
```


# Proportion of male and female dealates in pre-foundation population
```{r}
FigS3a <- read_excel("Dataset.xlsx", sheet="genotype of alates") %>% 
  as.data.frame(.) %>% 
  filter(stage == "pre_founding")
```

## * sample size
```{r}
summary_FigS3a <- FigS3a %>% 
  group_by(sex) %>% 
  summarize(number = n())
summary_FigS3a
```

## * figure
```{r}
Fig_S3a <-
ggplot(summary_FigS3a, aes(x="", y=number, fill=sex)) +
  geom_bar(width = 1, stat = "identity") + 
  coord_polar("y", start=0)
Fig_S3a

# save figure
# ggsave(file = "Figure S3a.pdf", plot = Fig_S3a, dpi = 300, width = 140, height = 89, units = c("mm"))
```

# Composition of founding units
```{r}
# data loading
FigS3b <- read_excel("Dataset.xlsx", sheet="composition of founding units") %>% 
  as.data.frame(.)
```

## * sample size
```{r}
summary_FigS3b <- FigS3b %>% 
  summarize(sum = sum(number_of_colony))
summary_FigS3b
```

## * figure
```{r}
FigS3b$combination <- factor(FigS3b$combination, levels=c("MMFF","MFFF","MMF","MFF","MM","MF","F","M"))

Fig_S3b <-
ggplot(FigS3b, aes(x="", y=number_of_colony, fill=combination)) +
  geom_bar(width = 1, stat = "identity") + 
  coord_polar("y", start=0)
Fig_S3b

# save figure
# ggsave(file = "Figure S3b.pdf", plot = Fig_S3b, dpi = 300, width = 140, height = 89, units = c("mm"))
```









